---
title: "Greenhouse gas emissions model"
author: "Cory and Eike"
bibliography:
 - references/packages.bib
 - references/papers.bib
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, include = F}
#install and load packages
library(decisionSupport)
library(ggplot2)
library(plyr)
library(rmarkdown)
# devtools::install_github("CWWhitney/uncertainty")
library(uncertainty)

#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'decisionSupport'), 'references/packages.bib')

```


## Generating the GHG emissions model

We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport [@R-decisionSupport], and to possibly optimize the model.

The formulas come from Chapter 10 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories [@dong_ipcc_2006]. The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).  Many variables are based on the work done in 'Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya' [@wilkes_variation_2020] and in 'Methods and guidance to support MRV of livestock emissions' [@wilkes_methods_2019].

*Here are our changes and additions to the original model:*

```{r load_data}

library(decisionSupport)
###CW Note ####
# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
###CW Note ####
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "data/livestock_ghg_input_table.csv"
```

```{r make_variables}
# Generate a function for testing the model function 'line by line' 
# by taking a single random sample of the provided estimates
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}

# run the function on the input_table
make_variables(estimate_read_csv(input_table))
```

```{r ghg_emissions_function}
# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

# Enteric fermentation emissions ($CH_4$)
  ####CW Note ####
  #####in these code sections I am missing the overview. 
  ##### it is not immediately clear what this is meant to do... 
  ##### from a specific publication? 
  
# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
#####EL CW Note ####
## We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.  
  
# We start with the net energy for maintenance (NE_m; in MJ/day)
# This includes a maintenance coefficient, for which different values are recommended depending on animal type
NE_m<-maintenaince_coefficient_Cfi*(live_weight_LW^0.75)

# The energy use for specific activities is calculated by multiplying NE_m by an activity coefficient (C_activity).

NE_activity<-C_activity*NE_m

#Net energy for growth (NE_growth; MJ/day):
# here, C is an animal type coefficient
# ### CW Note #### Assuming 'C' is C_activity or should it have been animal_type
C <- C_activity
# C <- animal_type
NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*
                   (0.96*SWG)^1.097)

# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)

##### EL CW Note ####
# SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.

##### EL CW Note #### 
#This used to be the following equation, which we didn't find a source for:
#NE_growth<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))

##### EL CW Note #### 
#The code doesn't include net energy mobilization by animals that lose weight (p. 314)

#Net energy for lactation

NE_lactation<-milk_yield*(1.47+0.40*milk_fat)

#variables:
#milk_yield (kg milk per day)
#milk_fat (percent)

#Draft animal services are not accounted for

#Energy need during pregnancy 
NE_pregnancy<-0.10*NE_m

##### EL CW Note #### 
#in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.

#We've now covered all equations up to eq 10 (p. 315)

##### EL CW Note #### 
#Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere

#Now we calculate the ratio between net energy consumed and the digestible energy consumed, for maintenance (NE_DE) and growth (NEg_DE) (eqs. 12 and 13)

NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+
            (1.126*(10^-5)*(dig_energy^2))-
              (25.4/dig_energy))

NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+
            (1.308*(10^-5)*(dig_energy^2))-
              (37.4/dig_energy))

##### EL CW Note #### 
#the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
NE_mobilized<-0
#same for extra energy spent by working animals
NE_w<-0

gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+
                          (NE_growth/NEg_DE))/(dig_energy/100)

enteric_CH4<-(gross_energy_intake*Y_m*365)/55.65 

##### EL CW Note #### 
######Y_m  here is the methane conversion rate -
###### Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
##### EL CW Note #### 
##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2<-enteric_CH4*25

# Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
# daily volatile solid excreted (page 10.42 of 2006 IPCC Guidelines for National Greenhouse Gas Inventories)

volsol<-(gross_energy_intake*
           (1-(dig_energy/100))+
           (urinary_energy*gross_energy_intake))*
             ((1-ash_content)/18.45) 

# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
# treat these as constants for each animal type. The following line assigns the corresponding Bo value
##### EL CW Note #### 
##### # to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.

Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))

# the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
mm_ch4<-(volsol*365)*
        (Bo*0.67*
           ((mms_pasture*MCFpasture)+
            (mms_spread*MCFspread)+
            (mms_drylot*MCFdrylot)+
            (mms_solidstore*MCFsolidstore)+
            (mms_composted*MCFcomposted)+
            (mms_liquid*MCFliquid)+
            (mms_biogas*MCFbiogas)+
            (mms_burn*MCFburn)+
              # the original code was lacking the 'sold' fraction - we added this for completeness
            (mms_sold*MCFsold)))

# Estimate CO2eq for manure methane
### ##### EL CW Note #### 
### this one also contained a multiplication by 365 that seemed erroneous
mm_CO2<-mm_ch4*25

##### #####
##### 
##### WE STOPPED HERE
#####
##### #####

# Calculate Nitrous oxide emissions from manure and urine on pasture
# ### ##### CW Note #### 
# ### ##### To get this to run I am assuming a temp 10% pregnancy_coefficient_Cp
pregnancy_coefficient_Cp <- 0.1
### #####
### ####
n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
n_retension<-((milk_yield*(3.3/100))/6.38)
n_excretion<-n_intake*(1-n_retension)*365
mm_direct_n2o<-n_excretion*((mms_pasture*0.02)+(mms_spread*0.005)+(mms_drylot*0.02)+(mms_solidstore*0.005)+
               (mms_composted*0.006)+(mms_liquid*0.005)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28)

# Nitrous oxide (N2O) from volatization 
mm_vol_n2o<-(n_excretion*((mms_pasture*0.02)+(mms_spread*0.07)+(mms_drylot*0.2)+(mms_solidstore*0.3)+
            (mms_composted*0.3)+(mms_liquid*0.4)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28))

# Nitrous oxide (N2O) from leaching 
mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

# Estimate CO2eq for manure nitrous
mm_CO2_N2Od<-mm_direct_n2o*298*365
mm_CO2_N2Ovol<-mm_vol_n2o*298*365
mm_CO2_N2Oleach<-mm_leach_n2o*298*365
mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol+mm_CO2_N2Oleach

# Emissions from feed production and transport. 
feed_CO2<-feed_prod_CO2*feed_trans_CO2

# Total emissions
on_farm<-enteric_CH4+enteric_CO2+mm_CO2+mm_CO2_N2O+feed_CO2

#CW Note ####
# Not clear what this part does (through to the ggplot)

# Relationship between greenhouse gas intensity and milk yields. 
# Fat adjusts
ann_adj <-kg_FPCM_year*365 
df <- data.frame(
  GHGi = on_farm/ann_adj, 
  FPCM = ann_adj)

# ### ##### CW Note
# Removing this plot
# plot

#library(ggplot2)
#ggplot(df, aes(x = FPCM, y = GHGi)) +geom_point() + ylim(0,50) 

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

```

Apply the `decisionSupport` tools for running the Monte Carlo [@R-decisionSupport]. See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.

```{r monte_carlo_simulation}

GHG_simulation <- mcSimulation(
  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
  model_function = ghg_emissions,
  numberOfModelRuns = 1e2, #100 for now
  functionSyntax = "plainNames"
)

```

```{r plot_distributions}

plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Methane Emissions from Enteric Fermentation")

```

# uncertainty

Here we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`. 

Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest {in_var} (x) and an outcome variable {out_var} (y) with associated and estimated densities with loess smooth linear fits density curves. Plot made with the scatter.hist() function in the plyr() package [@R-plyr] in R [@R-base]. 

Check the `milk_yield` effect on `enteric_CH4`.

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4)
```

Use `varkernel()` for another type of graphical representation of the same relationship. `varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable {out_var} and the influence variable of interest {in_var} using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002). The function produces a matrix of the estimated density (z) of the expected value of the outcome variable {out_var} (y) and the influence variable of interest {in_var} (x). It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a bivariate normal kernel.

Check the effect of {in_var} `milk_yield` on the {out_var} `enteric_CH4`.

```{r name_in_out_vars}
in_var  <- GHG_simulation$x$milk_yield
out_var <- GHG_simulation$y$enteric_CH4
```

```{r varkernel}
uncertainty::varkernel(in_var = in_var, out_var = out_var)
```

`varkernel()` shows a density surface plot of a given influence variable of interest {in_var} (x) and the outcome variable {out_var} (y). The legend shows the value for the estimated density (z). The plot is made with the `filled.contour()` function of the `graphics()` package.

In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable {out_var} along y-axis) given a specific value for a given influence variable of interest {in_var} (along x-axis). 

## Estimated outcome variable {out_var} given the expected variable

`varkernelslice()` calculates the estimated outcome variable {out_var} given the expected value of the influence variable of interest {in_var}, based on a slice of 'z' from the Kernel density calculated with `varkernel()`. The `expectedvariable` parameter is set to 30.

```{r varkernelslice}
uncertainty::varkernelslice(in_var = in_var, out_var = out_var, expectedin_var = 30)
```

`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis). Since this is a cut through the density kernel `varkernel()`, which integrates to 1, the probability values are relative, not absolute measures.

## Influencing variable intervals

`varviolin()` determines different possible influencing variable intervals {in_var} by calculating the optimal interval width for the given variable values. This is done using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).

$Os = 2 * IQRs / (obs_IQRs)^{(1/3)}$

Where $Os$ is The optimal interval width for our sample, $IQRs$ is the interquartile range for our sample and $obs_IQRs$ is the total number of observations for the interquartile range for our sample.

```{r varviolin}
uncertainty::varviolin(in_var = in_var, out_var = out_var)
```

`varviolin()` shows violin plots of influencing variable {in_var} values (x) and outcome variable {out_var} (y) values, with six different intervals of influencing variable {in_var} values. Plot made with `ggplot2()` [@R-ggplot2].

## Probability of outcome variable given variable

`varkernelslicerange()` shows the influencing variable {in_var} value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`. Here we set the maximum variable value to 4 and the minimum to 1.

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = 1, max_in_var = 4)
```

`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis). As with `varkernelslice()` the probability values shown are relative, not absolute measures. They are the result of cuts through the density kernel `varkernel()`, which integrates to 1. 

This document was generated using R Markdown [@R-rmarkdown].

# References